DATA 621 - Business Analytics and Data Mining

### 2020-11-20 HW5/HW5-Draft.Rmd
Built Poisson model. Need to explain it.

```{r graphsD, fig.height=5, fig.width=7}
# Using epanechnikov kernel
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnDUB], aes(x = value)) +
  geom_density(kernel = 'epanechnikov') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Kernel-smoothed Density of Numeric Predictors')
```

@@

```{r graphsH, fig.height=2, fig.width=7}
# Freedman-Diaconis rule for bin widths
FDbin <- function(x) {
  result <- 2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
  return(ifelse(result == 0, 0.5, result))
}
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnINT], aes(x = value)) +
  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Histogram of Integral Predictors')
```

@@

### Correlations
The corrgram below graphically represents the correlations between the numeric
predictor variables, when ignoring the missing variables.

```{r corrgram, fig.width=6, fig.height=4}
corrplot::corrplot(cor(DT[, ..nmtrn], use = 'complete.obs'),
         method = 'ellipse', type = 'lower', order = 'hclust',
         hclust.method = 'ward.D2')
```

@@

# Modeler # 1
## Missing Data
The missing daya

@@

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r statsN}
```
```{r graphsD}
```
```{r graphsH}
```
```{r corrgram}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r m2clean}
```
```{r m3clean}
```
```{r model1DummyFreq}
```
```{r model1RunFreq}
```
```{r model1DummySev}
```
```{r model1RunSev}
```
```{r model2RunFreq}
```
```{r model2RunSev}
```
```{r model3RunBaseLog}
```
```{r model3RunFreq}
```
```{r model3RunBaseLin}
```
```{r model3RunSev}
```
```{r setUpCompTables}
```
```{r validateM1F}
```
```{r validateM1S}
```
```{r validateM2F}
```
```{r validateM2S}
```
```{r validateM3F}
```
```{r validateM3S}
```
```{r freqComp}
```
```{r sevComp}
```
```{r badSev}
```
```{r evalProcessClean}
```
```{r evalPredictFreq}
```
```{r evalPredictSev}
```

>>>>

```{r graphsD, fig.height=5, fig.width=7}
# Using epanechnikov kernel to generate kernel-smoothed densities
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnDUB], aes(x = value)) +
  geom_density(kernel = 'epanechnikov') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Kernel-smoothed Density of Numeric Predictors')
```

@@

```{r graphsH, fig.height=2, fig.width=7}
# Freedman-Diaconis rule for bin widths
FDbin <- function(x) {
  result <- 2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
  return(ifelse(result == 0, 0.5, result))
}
# Generate histogram
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnINT], aes(x = value)) +
  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Histogram of Integral Predictors')
```

@@

## Correlations
The corrgram below graphically represents the correlations between the numeric
predictor variables, when ignoring the missing variables.

```{r corrgram, fig.width=5, fig.height=6.5}
# Create corrgram
corrplot::corrplot(cor(DT[, ..nmtrn], use = 'complete.obs'),
         method = 'ellipse', type = 'lower', order = 'hclust',
         hclust.method = 'ward.D2')
```

@@

## Modeler # 1
### Missing Data
The missing data can be split into two groups: `STARS` and the rest. For all the
others, I do not think there is signal encoded in the "missingness" and thus
will use some form of imputation. `STARS` is different as discussed in the DATA
EXPLORATION section. Therefore, I will create a new feature called `Rated` which
will be true if `STARS` is missing, and then only use this variable and its
interaction with `STARS` in the model, and not `STARS` by itself.

```{r model1addVars}
# Copy the training set to leave a pristine version for others
m1trnX <- copy(trnX)

# Add the "rated" factor variable
m1trnX[, Rated := factor(ifelse(is.na(STARS), 'Unrated', 'Rated'),
                         levels = c('Unrated', 'Rated'),
                         labels = c('Unrated', 'Rated'))]
```

### Imputation
Imputation will be handled through bagging. Instead of looking at a nearest
neighbors approach, which also requires centering and scaling, once can create a
set of bagged trees. As per Kuhn (2019):

>*For each predictor in the data, a bagged tree is created using all of the*
*other predictors in the training set. When a new sample has a missing*
*predictor value, the bagged model is used to predict the value. While, in*
*theory, this is a more powerful method of imputing, the computational costs*
*are much higher than the nearest neighbor technique.*

The bagged trees will use `STARS` as a predictor, as that contains valuable
information, but prior to the imputation, the missing `STARS` values will be
replaced by 0 so as not to be imputed, as per the explanation above. The
preprocessing will also check for near-zero value predictors.

```{r model1impute}
# Create bagged-tree imputation model
m1trnXi <- preProcess(m1trnX, method =c('nzv', 'bagImpute'))

# Replace missing STARS with 0s
m1trnX[is.na(STARS), STARS := 0L]

# Impute the remaining NAs
m1trnXimp <- predict(m1trnXi, newdata = m1trnX)
```

It will be instructive to compare the shapes of the distributions after
imputation to those before imputation.

```{r m1postImputeD, fig.height=5, fig.width=7}
# Isolate numeric only predictors
m1predictorDT <- m1trnXimp[, .SD, .SDcols = nmtrn]

# Melt them from wide to long format
m1predictorDTM <- melt(m1predictorDT, variable.name = 'metric',
                     value.name = 'value',  variable.factor = FALSE,
                     id.vars = NULL)

# Using epanechnikov kernel to generate kernel-smoothed densities
ggplot(predictorDTM[metric %chin% nmtrnDUB], aes(x = value)) +
  geom_density(kernel = 'epanechnikov') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Kernel-smoothed Density of Numeric Predictors')
```

Thankfully, none of the shapes appears to have changed significantly, which
implies the imputation was in line with the original distributions. Now for the
histograms.

```{r m1postImputeH, fig.height=2, fig.width=7}
# Generate histogram
ggplot(m1predictorDTM[metric %chin% nmtrnINT], aes(x = value)) +
  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Histogram of Integral Predictors')
```

Here too, `AcidIndex` and `LabelAppeal` look very similar to their
pre-imputation distribution. What stands out is `STARS`. Both the Poisson and
the negative binomial models are either unimodal or have to consecutive values
sharing the mode, due to their inherent nature as discrete disctibutions. What
is seen here is that 1 star is empirically less common than 0 (unrated) or 2
stars. On the one hand, this could simply be noise (process variance) or
parameter error (finite sample size). On the other, it could indicate the need
for a zero-inflated version of the counting distributions.

# BUILD MODELS
## Modeler # 1
Each of the models will follow a similar format, with only the family changing.
Each model will start with the saturated model including all the predictors,
***except*** for `STARS` which will be replaced with the interaction between
`Rated` and `STARS`. A forward and backwards stepwise procedure based on AIC
will be used, and the model with the lowest AIC on the training set will be the
selected to be tested against the testing set. Using stepAIC precludes the use
of k-fold cross validation.

```{r m1TrainControl}
# All Modeler 1 models will use stepAIC so turn off cross-validation
trc <- trainControl(method = 'none')
```

### Poisson Model

```{r m1Poisson}
set.seed(350047)
m1PFit <- train(TARGET ~ . - INDEX - STARS + Rated:STARS, data = m1trnXimp,
                trControl = trc, method = 'glmStepAIC',
                family = poisson(link = 'log'), direction = 'both', trace = 0)
# Print results
kable(summary(m1PFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 1 Poisson Regression Output")
```

```
                     Df Deviance   AIC
<none>                    9654.5 31998
- Sulphates           1   9657.1 31998
+ Density             1   9654.1 31999
+ ResidualSugar       1   9654.2 31999
- Alcohol             1   9658.3 32000
+ CitricAcid          1   9654.4 32000
+ FixedAcidity        1   9654.4 32000
- pH                  1   9659.0 32000
- Chlorides           1   9659.5 32001
- FreeSulfurDioxide   1   9659.7 32001
- TotalSulfurDioxide  1   9670.3 32012
- VolatileAcidity     1   9670.9 32012
- AcidIndex           1   9880.9 32222
- LabelAppeal         1  10102.7 32444
- `STARS:RatedRated`  1  10335.1 32676
- RatedRated          1  10337.5 32679
```

@@

# References

  * Kuhn, M. (2019, March 27). *The `caret` Package*.
  https://topepo.github.io/caret/pre-processing.html#imputation

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r statsN}
```
```{r graphsD}
```
```{r graphsH}
```
```{r corrgram}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r model1impute}
```
```{r m1postImputeD}
```
```{r m1postImputeH}
```
```{r m1TrainControl}
```
